---
title: "NHTSA DSFS Meta-Analysis"
author: "Dan Flynn"
date: "September 27, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

setwd("~/git/dsfs")

source("MA_Analysis_Prep.R")
```

### Background

Quantitative meta-analysis of the effectiveness of Dynamic Speed Feedback Signs (DSFS) in reducing driving speeds.

Hypotheses:

Hypothesis        | Calculation    
------------------|------------------------------------------------------------
H1A. Activation   | (Upstream - Adjacent)_Before - (Upstream - Adjacent)_During
H1B. Activation   | Before - During
H2A. Deactivation | (Upstream - Adjacent)_During - (Upstream - Adjacent)_After
H2B. Deactivation | During - After



Articles were located from searches the databases TRID, National Transportation Library (NTL), WorldShare, Academic Search Complete, PsycINFO, Web of Science, and Science Direct. The searches were done between 3/9/2016 and 3/16/2016.  Any of the search terms in the tables that ended in 's' were wildcards with *.

## Activation hypothesis

```{r, echo=FALSE}


```


